---
title: "Lab3A_jeanniecarreiro"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# dont forget to call the library each time

```{r}
library(tidyverse)
```

# had to made a folder called data to run the code below which send the data into the folder

```{r}
 download.file(url="https://ndownloader.figshare.com/files/2292169",
              destfile = "data/portal_data_joined.csv")
```

# now, we load the data and check the data

```{r}
surveys <- read_csv("data/portal_data_joined.csv")
surveys
head(surveys)
view(surveys)
```

# inspect the dataframe and its objects
# https://datacarpentry.org/R-ecology-lesson/02-starting-with-data.html go to this link if you forget what they do

```{r}
str(surveys)
dim(surveys)
nrow(surveys)
ncol(surveys)
head(surveys)
tail(surveys)
names(surveys)
data.frame
rownames(surveys)
str(surveys)
summary(surveys)
```

Challenge questions (set 1)
The class of the object surveys is a data.frame
34,786 rows by 13 columns
There are 48 species

# indexing and subsetting data frames to extract specific data you need to specify rows and columns

# here are the different ways to do that that end up with the same result

```{r}
# first element first column, vector
surveys[1, 1]
# first element 6th column, vector
surveys[1, 6]
# first column, vector
surveys[,1]
# first column, data.frame
surveys[1]
# first 3 elements in 7th column, vector
surveys[1:3, 7] 
# 3rd row, data.frame
surveys[3, ]
# same as head(surveys)
head_surveys <- surveys[1:6 ] 
```

# : is a function that created numeric vectors of integers in increasing order 

```{r}
10:1
```

# exclude indices of a data frame using -

```{r}
# except the first column
surveys[, -1]
# same as head(surveys)
surveys[-(7:34786), ]
# subset data frames by calling column names (below)
# data.frame (1)
surveys["species_id"] 
# vectors (3)
surveys[, "species_id"]
surveys[["species_id"]]
surveys$species_id
```

Challenge questions (set 2)
Create a data.frame (surveys_200) containing only the data in row 200 of the surveys dataset.

Notice how nrow() gave you the number of rows in a data.frame?

Use that number to pull out just that last row in the data frame.
Compare that with what you see as the last row using tail() to make sure it’s meeting expectations.
Pull out that last row using nrow() instead of the row number.
Create a new data frame (surveys_last) from that last row.
Use nrow() to extract the row that is in the middle of the data frame. Store the content of this row in an object named surveys_middle.

Combine nrow() with the - notation above to reproduce the behavior of head(surveys), keeping just the first through 6th rows of the surveys dataset.

```{r}
# Challenge questions set 2
surveys200 <- surveys[200, ]
nrows <- nrow(surveys)
surveyslast <- surveys[nrows, ]
surveysmiddle <- surveys[nrows/2, ]
surveyshead <- surveys[-(7:nrows), ]
```

